---
title: "basic_raster_data_functions"
author: "Duke"
date: "2022-08-18"
output: html_document
---

## About Raster data

The data metadata about raster can be accessed using various functions.

The most basic ones are...

-   `extent()` - shows the spatial extent

-   `crs()` - shows the coordinate reference system

-   `ncell()` - shows the number of grid cells

```{r}
# load the raster package
library(raster)

# read in the rasters
canopy <- raster("data/canopy.tif")
manhatttan <- brick("data/manhattan.tif")

# get the extent of canopy layer
extent(canopy)
crs(manhattan)

ncell(canopy)
ncell(manhattan)
```

### Accessing raster data values

to show that raster is in memory, one can use `inMemory()` function on an object

If you use the `head()` function, the raster package will read in only the values needed. not the full set of values

If you perform spatial analysis that require it, you can read in in values from a raster manually with the function `getValues()`

```{r}
# determine if canopy has been read in
inMemory(canopy)

head(canopy)

#use getValues to read in the values into a vector
vals <- getValues(canopy)

# use hist to create a quick histogram of the values
hist(vals)
```

### Plot raster object

one can use `plot()` function for single band or `plotRGB()` to plot the raster layers together as a single image

```{r}
plot(canopy)
```

```{r}
plot(manhattan)
```

```{r}
plotRGB(manhattan)
```

### Vector and raster coordinate systems

Spatial Operations on multi-layered datasets require that the datasets have the same coordinate reference systems. use `st_crs()` on vector data and `crs` on the raster crs can be defined using espg or proj4string

```{r}
# determine the crs for neighborhoods and trees
st_crs(neighborhoods)
st_crs(trees)
```

```{r}
# Assign crs for trees
csr_1 <- "+proj=longlat +ellps=WGS84 +no_defs"
st_crs(trees) <- csr_1

```

```{r}
# determine the CRS for the canopy
crs(canopy)
crs(manhattan)
```

```{r}
# assing the crs for manhattan
crs_2 <- "+proj=utm +zone=18 +ellps=GRS80 +datum=NAD83 +units=m +no_defs"
crs(manhattan) <- crs_2
```

### Transforming your layers to a common crs

The process of transforming crs of layers to have the same crs for purposes of spatial analysis is called projection.

To check if your object has a projected crs crs, you can look at the part of the result `st_crs()` or `crs`() **If it begins with +proj=longlat then your crs is unprojected**

For raster, use `projectRaster()` to prevent distortion.

```{r}
the_crs <- st_crs(canopy, asText=TRUE)

# project trees to match the CRS canopy
trees_crs <- st_transform(trees, crs=the_crs)

# project neighborhoods to match the crs for canopy
naighborhood_crs <- st_transform(neighborhoods, crs=the_crs)

#project manhattan to math the crs of canopy
manhattan_crs <- projectRaster(manhattan, crs=the_crs, method="ngb")

st_crs(trees_crs)
st_crs(neighborhoods_crs)
crs(manhattan_crs)
```

### Plot vector and raster data together

```{r}
plot(canopy)
plot(neighborhoods, add=TRUE)
```

```{r}
crs(neighborhoods)
crs(canopy)
```

Neighborhoods cant be seen because the crs are different

do a reprojection

```{r}
neighborhoods_crs <- st_transform(neighborhoods, crs = the_crs)
plot(canopy)

plot(neighborhoods_crs, add=TRUE)

tm_shape(canopy) + 
    tm_raster() + 
    tm_shape(neighborhoods_crs) + 
    tm_polygons(alpha = 0.5)
```

### Dropping geometry from a dataframe

if you want to force the geometry to b dropped, use the fs function `st_set_geometry()` and set the geometry to be null

```{r}
# create a dataframe of counts by species
species_counts = count(trees, species, sort = TRUE)

head(species_counts)

# drop the geometry column

st_set_geometry(species_counts, NULL)

head(species_counts)
```

## Join spatial and non spatial data

```{r}
# limits the fields boro_name, county_fip, boro_code

boro <- select(neighborhoods, boro_name,county_fip, boro_code)

# drop the geometry column
boro_no_geometry <- st_set_geometry(boro, NULL)

# get only distinct countries
boro_distinct <- distinct(boro_no_geometry)

# join the country detail into trees object
trees_with_county <- inner_join(trees, boro_distinct, by =c("boroname" = "boro_name"))

head(trees_with_county)
```

### Simplify the neighborhoods boundaries

One can use the `st_simplify()` to reduce line and polygon complexity

```{r}
library(pryr)
# plot geometry of neighborhoods
plot(st_geometry(neighborhoods), col="grey")

# measure the size of the neighborhoods object
object_size(neighborhoods)

# compute the number of vertices in the neughborhoods object

pts_neighborhoods <- st_cast(neighborhoods$geometry, "MULTIPOINT")
cnt_neighborhoods <- sapply(pts_neighborhoods, length)
sum(cnt_neighborhoods)

# simplify the number of vertices in the neighborhoods object

neighborhoods_simple <- st_simplify(neighborhoods, preserveTopology = TRUE, dTolerance = 100)

# compute the object size of neighborhoods simple
object_size(neighborhoods_simple)
pts_neighborhoods_simple <- st_cast(neighborhoods_simple$geometry, "MULTIPOINT")
cnt_neighborhoods_simple = sapply(pts_neighborhoods_simple, length)
sum(cnt_neighborhoods_simple)

plot(st_geometry(neighborhoods_simple), col="grey")
```

### Converting objects to sp objects

```{r}
# read in trees shapefile 
trees <- st_read("data/trees/trees.shp")

# convert to Spatial class
trees_sp <- as(trees, Class = "Spatial")

class(trees_sp)

trees_sf <- st_as_sf(trees_sp)
class(trees_sf)
```

### Converting to and from cooordinates

```{r}
# read in trees scv

trees_df <- read.csv("data/trees.csv")

# convert the dataframe into an sf object
trees_sf <- st_as_sf(trees_df, coords = c("X", "Y"), crs=4326)
plot(st_geometry(trees_sf))

# write the file out with coordinates

st_write(trees_sf, "new_trees.csv", layer_options = "GEOMETRY=AS_XY", delete_dsn = TRUE)

# read in the file
new_trees = read.csv("new_trees.csv")
head(new_trees)
```
