---
editor_options: 
  markdown: 
    wrap: 72
---

## Reading data

**Vector data** - use `sf` package and `st_read()` function

**Raster data** - use the raster package and `raster()` or `brick()`
function

`raster()` function is for reading single band rasters `brick()`
function is for reading multi-band rasters

## Writing data

**Vector data** - use st_write(m_poly, "path/to/file.ext")

**Raster data** - writeRaster(my_raster, "path/to/raster.ext")

```{r}
# load sf package
library(sf)

# read in the trees shapefile
trees <- st_read("data/trees/trees.shp")

# read in neighborhoods
neighborhoods <- st_read("data/neighborhoods/neighborhoods.shp")

# read in parks
parks <- st_read("data/parks/parks.shp")

# view the first few trees
head(trees)
```

### Reading Raster data

```{r}
# load library raster
library(raster)

# read in tree camopy single-band raster
canopy <- raster("data/canopy.tif")

# Read in manhattan Landsat image multi-band raster
manhattan <- brick("data/manhattan.tif")

# ge the class for the new objects
class(canopy)
class(manhattan)
```

#### sf objects and data frames

sf are just data frames with some special properties. this means that
packages like \`dplyr\` can be used to manipulate sf objects

```{r}
library(dplyr)

# use filter to limit honey locust trees
honeylocust <- trees %>% filter(species == "honeylocust")

# count the number of rows
nrow(honeylocust)

# limit to tree_id and boroname variables

honeylocust_lim <- honeylocust %>% select(tree_id, boroname)
head(honeylocust_lim)
```

### Geometry as list-columns in sf

A list-column behaves to a certain extent like any other R column. the
main difference is that instead of a standard value such as a single
number, character or boolean value, each observation value in that
column is a piece of an R list and that list can be as complex as
needed. The list-column allows you to store farmore information in a
single variable. so sf takes this advantage to store all geographic
information for each feature in the list.

```{r}
library(tibble)
# create a sampledataframe with one column
df <- data.frame(a=1:3)

df

# add a list column to the data frame
df$b <- list(1:4, 1:5, 1:10)

df

# convert dataframe to a tibble
as_tibble(df)

# pull out the third observation from both columns
df$a[3]
df$b[3]
```

## Extracting geometric information from vector layers

```{r}
# if not loaded, read in parks shapefile

# compute areas of the parks

# incase of an error, turn off spherical geometry
sf_use_s2(FALSE)
areas <- st_area(parks)

# cerate a quick histogram of the areas usin g hist
hist(areas, xlim = c(0, 200000), breaks=1000)

# filter the parcks greater than 3000(square meters)
head(parks)
```

Note that the result of functions like `st_area()` and `st_length()`
will not be a traditional vector. Instead, the result has a class of
units. which means the vector result is accompanied by metadata
describing the object's units.

So you need to remove the units with `unclass()`

or convert the val's class to units

```{r}
# filter to parks greater than 3000 (square meters)
big_parks <- parks %>% filter(unclass(areas) > 3000)
plot(st_geometry(big_parks))
```
