---
title: "Geoprocessing with rasters"
author: "Duke"
date: "2022-08-23"
output: html_document
---

### Limit rasters to focus areas

With `mask()` you essentially place your ares of interest on top of the raster and any raster cells outside the boundary are assigned NA values

```{r}
library(sf)
library(raster)
parks <- st_read("data/parks/parks.shp")
canopy <- raster("data/canopy.tif")

parks_cp <- st_transform(parks, crs = crs(canopy, asText=TRUE))

# compute the areas of the parks
areas <- st_area(parks_cp)
# filter the parks with areas > 3000

parks_big <- filter(parks_cp, unclass(areas) > 3000)

plot(canopy)

plot(st_geometry(parks_big))

# convert parks to spatial object
parks_sp <- as(parks_big, "Spatial")

canopy_mask <- mask(canopy, mask=parks_sp)
```

### Cropping

```{r}
# Convert the parks_big to a Spatial object
parks_sp <- as(parks_big, "Spatial")

# Mask the canopy with the large parks 
canopy_mask <- mask(canopy, mask = parks_sp)

# Plot the mask
plot(canopy_mask)

# Crop canopy with parks_sp
canopy_crop <- crop(canopy, parks_sp)

# Plot the cropped version and compare
plot(canopy_crop)
```

### Extracting values from raster

```{r}
# Project the landmarks to match canopy
landmarks_cp <- st_tranform(landmarks, crs = crs(canopy, asText = TRUE))

# Convert the landmarks to a Spatial object
landmarks_sp <- as(landmarks_cp, "Spatial")

# Extract the canopy values at the landmarks
landmarks_ex <- extract(canopy, landmarks_sp)

# Look at the landmarks and extraction results
landmarks_cp
landmarks_ex
```

### Raster math with overlay

Using the imperviousness layer from the USGS. Imperviousness measures whether water can pass through a surface. So a high percentage impervious surface might be a road that does not let water pass through while a low percentage impervious might be something like a lawn.

We will identify the **most** urban locations by finding areas that have both a low percentage of tree canopy ([removed] 80%). To do this, we defined the function `f` to do the raster math for you.

```{r}
# read in data
impervious <- raster("data/impervious.tif")

# function f with 2 args and the raster math code

f <- function(rast1, rast2) {
  rast1 < 20 & rast2 > 80
}

# do the overlay using f as fun

canopy_imperv_overlay <- overlay(canopy, impervious, fun = f)

plot(canopy_imperv_overlay)
```
