---
title: "Manipulating Raster Layers"
author: "Duke"
date: "2022-08-22"
output: html_document
---

## Reclassification and Aggregation

to reduce the resolution of raster - use `aggregate()` function which aggregates grid cells into larger grid cells using user defined function like mean, max or mode amount if aggregations ins driven by the fact arg, default is 2

```{r}
library(raster)
canopy <- raster("data/canopy.tif")
plot(canopy)

# determine the resolution of the canopy
res(canopy)

# detemine the number of cells
ncell(canopy)

# aggregate he raster
 canopy_small <- aggregate(canopy, fact=10)
 
 plot(canopy_small)
 res(canopy_small)
 ncell(canopy_small)
```

### Change values and handle missing values in rasters

```{r}
vals <- cbind(100, 300, NA)

# reclassify 
canopy_reclass <- reclassify(canopy, rcl = vals)

plot(canopy_reclass)
```

```{r}
library(sf)
manhattan <- brick("data/manhattan.tif")
# Review df
df

# Convert the data frame to an sf object             
df_sf <- st_as_sf(df, coords = c("longitude", "latitude"), crs=4326)

# Transform the points to match the manhattan CRS
df_crs <- st_transform(df_sf, crs = crs(manhattan, asText = TRUE))

# Buffer the points
df_buf <- st_buffer(df_crs, dist = 1000)

# Plot the manhattan image (it is multi-band)
plotRGB(manhattan)
plot(st_geometry(df_buf), col = "firebrick", add = TRUE)
plot(st_geometry(df_crs), pch = 16, add = TRUE)
```

```{r}
# Read in the neighborhods shapefile
neighborhoods <- st_read("data/neighborhoods/neighborhoods.shp")

# Project neighborhoods to match manhattan
neighborhoods_tf <- st_transform(neighborhoods, crs = 32618)

# Compute the neighborhood centroids
centroids <- st_centroid(neighborhoods_tf)

# Plot the neighborhood geometry
plot(st_geometry(neighborhoods_tf), col = "grey", border = "white")
plot(centroids, pch = 16, col = "firebrick", add = TRUE)
```

### Bounding Boxes, dissolve features and create a convex hull

```{r}
plot(neighborhoods, col="grey", border="white")
plot(beech, add=TRUE, pch=16,col="forestgreen")
st_bbox(beech)

beech_bbox <- st_make_grid(beech, n=1)
plot(st_geometry(neighborhoods), col="grey", border="white")
plot(beech, add=TRUE, pch=16, col="forestgreen")
plot(beech_bbox, add=TRUE)
```

### Dissolve multiple features into One

If you have a set of points and you want to draw a convex hull around them you first need to bundle the points into a single MULTIPOINT feature and in order to do this, you will use the dissolve function in `sf` called `st_unioin()`

With polygons, `st_union()` will dissolve all the polygons into a single polygon representing the area where all the polygons overlap.

```{r}
# buffer the beach object by 3000
beech_buffer <- st_buffer(beech, 3000)

# Limit the object to just geometry
beech_buffers <- st_geometry(beech_buffer)

# compute the number of features in beech_buffers
length(beech_buffers)

#plot the tree buffers
plot(beech_buffers)

beech_buf_union <- st_union(beech_buffers)

length(beech_buf_union)

plot(beech_buf_union)
```

### Compute a convex hull around vectors

A more precise bounding polygon is sometimes needed, one that fits your data more tightly.

For this. you can use the `st_convex_hull()` function.

```{r}
head(beech)

beech1 <- st_union(beech)

head(beech1)

length(beech)
length(beech1)

# compute a tight bounding box
beech_hull <- st_convex_hull(beech1)

plot(beech_hull, col="red")
plot(beech1, add=TRUE)
```
