---
title: "Multilayered geoprocessing and relationships"
author: "Duke"
date: "2022-08-23"
output: html_document
---

### Spatial Joins

If you want to know how many trees are in each neighborhood but you dont have a neighborhood attribute in the tree data. the best way to do this is to use `st_join()`

`st_join()` function requires sf dataframes as input and will not accept object that is just sf geometry. You can use `st_sf()` function to convert sf geometry objects to an sf dataframe.

```{r}
plot(st_geometry(neighborhoods))

# plot the beech on top of the neighborhoods
plot(beech, add=TRUE, pch=16, col="red")

# convert the beech geometry to a sf dataframe
beech_df <- st_sf(beech)
head(beech_df)

# join the beech trees with the neighborhoods
beech_neigh <- st_join(beech_df, neighborhoods)
head(beech_neigh)
```

### Spatial relationships

```{r}
# Identify neighborhoods that intersect with the buffer
neighborhoods_int <- st_intersects(buf, neighborhoods)

# Identify neighborhoods contained by the buffer
neighborhoods_cont <- st_contains(buf, neighborhoods)

# Get the indexes of which neighborhoods intersect
# and are contained by the buffer
int <- neighborhoods_int[[1]]
cont <- neighborhoods_cont[[1]]

# Get the names of the names of neighborhoods in buffer
neighborhoods$ntaname[int]

# Clip the neighborhood layer by the buffer (ignore the warning)
neighborhoods_clip <- st_intersection(buf, neighborhoods)

# Plot the geometry of the clipped neighborhoods
plot(st_geometry(neighborhoods_clip), col = "red")
plot(neighborhoods[cont,], add = TRUE, col = "yellow")
```

### Measuring distance between features

```{r}
# Read in the parks object
parks <- st_read("data/parks/parks.sh")

# Test whether the CRS match
crs(empire_state) == crs(parks)

# Project parks to match empire state
parks_es <- st_transform(parks, crs = crs(empire_state))

# Compute the distance between empire_state and parks_es
d <- st_distance(empire_state, parks_es)

# Take a quick look at the result
head(d)

# Find the index of the nearest park
nearest <- which.min(d)

# Identify the park that is nearest
parks_es[nearest, ]
```
